Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-05-10 code/rmarkdown/article.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56). The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %). Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

>>>>

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56). The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($> 80$ %). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`; maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($> 95$ %). Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

### 2016-05-10 code/rmarkdown/figures.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation secured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	geom_text(x=0.05, y=0.95, label='(a)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation secured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	geom_text(x=0.05, y=0.95, label='(b)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	data.frame()

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	geom_text(x=0.05, y=0.95, label='(a)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	geom_text(x=0.05, y=0.95, label='(b)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	data.frame()

### 2016-05-10 code/rmarkdown/supporting_information.Rmd
+ manuscript fixes
Former-commit-id: 88df6830484ad38e28a669bf3a5581b30a3eba7c

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue. Note that \\textit{',gsub('\\_', '\\ ', missing.species),'} was omitted from analysis.')}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', face='italic'),
		axis.title=element_blank()
	) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) +
	facet_wrap(~ species, ncol=4)
```

@@

```{r, echo=FALSE}
plot.spp.nmds <- function(i) {
	## prepare data
	# get vector of adaptive and neutral column names
	adapt.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	neutral.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	# skip species if not used in analysis
	curr.FPLY <- grid.FPLY[,
		c(adapt.col, neutral.col,
		'long','lat','group'),drop=FALSE]
	curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
		d=gsub('_', ' ', d, fixed=TRUE)
	)
	for (j in seq_len(length(adapt.col)+length(neutral.col)))
		curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
	curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
	# make plot
	curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey85', color='grey70') +
			geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
				group='group', fill='value'), color='grey10') +
			theme_classic() +
			theme(
				axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
				panel.border=element_rect(color='black', fill=NA, size=1),
				strip.background = element_rect(fill='grey20'),
				strip.text = element_text(color='white', size=20)
			) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0),
				ylim=buffered.range(grid.FPLY$lat, 0)
			) +
			scale_fill_gradientn(name='',
				colors=brewer.pal(9, 'RdYlBu'),
				guide=guide_colorbar(ticks=element_line(color='black'),
					border=element_line(color='black'), barheight=unit(7, 'cm'))
			) + facet_wrap(~d)
	})
	do.call(grid.arrange, append(curr.GROBS, list(ncol=4)))
	invisible()
}

calc.fig.height <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	return((((n.plots-1) %/% 4)+1) * 2.5)
}

>>>>

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue. Note that \\textit{',gsub('\\_', '\\ ', missing.species),'} was omitted from analysis.')}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
) %>% mutate(presence=factor(c('absent', 'present')[presence+1]))

# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', face='italic'),
		axis.title=element_blank()
	) +
		scale_fill_manual(name='Occupancy', values=c('absent'='#FC8D59', 'present'='#91CF60')) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) +
	facet_wrap(~ species, ncol=4)
```

@@

```{r, echo=FALSE}
plot.spp.nmds <- function(i) {
	## prepare data
	# get vector of adaptive and neutral column names
	adapt.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	neutral.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	# skip species if not used in analysis
	curr.FPLY <- grid.FPLY[,
		c(adapt.col, neutral.col,
		'long','lat','group'),drop=FALSE]
	curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
		d=gsub('_', ' ', d, fixed=TRUE)
	)
	for (j in seq_len(length(adapt.col)+length(neutral.col)))
		curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
	curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
	# make plot
	curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey85', color='grey70') +
			geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
				group='group', fill='value'), color='grey10') +
			theme_classic() +
			theme(
				axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
				panel.border=element_rect(color='black', fill=NA, size=1),
				strip.background = element_rect(fill='grey20'),
				strip.text = element_text(color='white', size=20)
			) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0),
				ylim=buffered.range(grid.FPLY$lat, 0)
			) +
			scale_fill_gradientn(name='',
				colors=brewer.pal(9, 'RdYlBu'),
				guide=guide_colorbar(ticks=element_line(color='black'),
					border=element_line(color='black'), barheight=unit(7, 'cm'))
			) + facet_wrap(~d)
	})
	do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
	invisible()
}

calc.fig.height <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	return((((n.plots-1) %/% 4)+1) * 5)
}
